---
title: "GroupS_Milestone3"
author: "PhuongTseng"
date: '2022-11-07'
output:
  html_document:
    df_print: paged
---
# Clean dataset with descriptive statistics for relevant data elements

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
path <- "/home/rstudio/PHW251_Fall2022/phw251"
setwd(path)

library(readr)
library(data.table)
library(tidyverse)

#import demographic data
ca_county_demographic <- fread("./ca_county_demographic.csv")
#import mortality data
ca_county_mortality <- fread("./ca_county_mortality.csv")
#import healthcare construction data with fread
hcai_healthcare_construction <- fread("./hcai_healthcare_construction.csv")


```

#### Document the import process
```{r import, echo=TRUE}
#use dplyr::select specific columns
demog <-
  ca_county_demographic %>% 
  dplyr::select(name, pop2012, pop12_sqmi, med_age, owner_occ, renter_occ)

#check first 2 records
head(demog, 2)

#use dplyr::rename_with to make column names lower case
mortality <-
  ca_county_mortality %>% 
  dplyr::rename_with( ~ tolower(gsub(" ", "_", .x, fixed = TRUE)))

head(mortality, 2)

#same as above and substitute a dash "_" for space
construction <-
  hcai_healthcare_construction %>% 
  dplyr::rename_with( ~ tolower(gsub(" ", "_", .x, fixed = TRUE)))
#see the last 2 records
tail(construction, 2)
```

#### Subset rows or columns, as needed
- Create new variables needed for analysis (minimum 2)
  - New variables should be created based on existing columns; for example
    - Calculating a rate
    - Combining character strings
  - If no new values are needed for final tables/graphs, please create 2 new variables anyway
  
Using our 3 vectors:  
1) ca_county_demographic  
2) ca_county_mortality  
3) hcai_healthcare_construction  
  
```{r subset, eval=FALSE, include=FALSE}

ca_county_demographic

ca_county_mortality

hcai_healthcare_construction



```

#### Clean variables needed for analysis (minimum 2)
  - Examples
    - Recode invalid values
    - Handle missing fields
    - Recode categories
  - If not needed for final analysis, please create at least 2 new variables anyway
  
```{r}


```

#### Data dictionary based on clean dataset (minimum 4 data elements), including:
  - Variable name
  - Data type
  - Description
  
```{r}


```

#### One or more tables with descriptive statistics for 4 data element
```{r}


```

#### PDF that is professionally prepared for presentation
  - Each part of the milestone is clearly on one page (use \newpage to push to a new page)
  - Only the necessary information is outputted (you should suppress, for example, entire data frame outputs)
  - Use of headers and sub headers to create an organized document

```{r}


```

