---
title: "Group S: Milestone #2"
author: "PhuongTseng & Carri Beshears"
date: '2022-10-02'
output: pdf_document
---

This is a team assignment; each team should complete and turn in a PDF created 
from an Rmd via Github. Please include code and output for the following components:

## Description of dataset
What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
How does the dataset relate to the group problem statement and question?

ANSWER: This project will use 3 datasets to identify 5 California counties for additional funding for healthcare facility improvementent. First, information from the 2012 California county census will be reviewed to to isolate counties in rural areas, with residents of a high median age, and with a high proportion of renters over homeowners. Next, mortaility surviellance data will be used to determine mortality from chronic conditions will be aggregated by county for the years 2014-2020. Finally, HCAI data will be used to account for current healthcare funding by county from 2014-2020. All of these data sets are avaliable from the CA.gov open data portal. 

## Import statement
NOTE: Please use datasets available in the PHW251 Project Data github repoLinks to an external site. 
(This is important to make sure everyone is using the same datasets)

Use appropriate import function and package based on the type of file
Utilize function arguments to control relevant components 
(i.e. change column types, column names, missing values, etc.)

#### Document the import process
```{r import, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
getwd()
ca_county_demographic <- fread("/home/rstudio/PHW251_Fall2022/phuong/phw251/ca_county_demographic.csv")
ca_county_mortality <- fread("/home/rstudio/PHW251_Fall2022/phuong/phw251/ca_county_mortality.csv")
hcai_healthcare_construction <- fread("/home/rstudio/PHW251_Fall2022/phuong/phw251/hcai_healthcare_construction.csv")

head(ca_county_demographic, 2)
head(ca_county_mortality, 2)
tail(hcai_healthcare_construction, 2)

#Make hcai colnames lower case with gsub

#reshape ca_county_mortality then merge these two datasets

# data <- merge(ca_county_demographic, ca_county_mortality)

#remove spaces and column names to_lower for hcai
#split the county column and remove spaces and "-"


#  data2 %>% 
#  dplyr::rename(location = `Location information`) %>%
#  dplyr::rename_with(~ tolower(gsub(" ", "_", .x, fixed = TRUE))) %>% 
#  dplyr::mutate(
#    measure = if_else(
#      location %in% c("Colorado", "New York", "Hawaii"),
#      "imperial",
#      "metric"
#    ), 
#    bmi = if_else(
#      measure == "imperial", 
#      (703 * data2$weight) / (data2$height ^ 2), 
#      data2$weight / (data2$height ^ 2)
#  ),
#  bmi_var = case_when(
#      bmi < 18.5 ~ "underweight",
#      bmi < 24.9 & bmi >= 18.5 ~ "normal",
#      bmi < 29.9 & bmi >= 25.0 ~ "overweight",
#      bmi >= 30 ~ "obese"
#    )
#  ) %>%
#  arrange(desc(bmi)) #%>% select(location, gender)

# head(t, 3)

```

\newpage

#### Identify data types for 5+ data elements/columns/variables
- Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
- Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
- Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type?
```{r data elements, include=FALSE}




```

\newpage

#### Provide a basic description of the 5+ data elements
- Numeric: mean, median, range
- Character:  unique values/categories
- Or any other descriptive that will be useful to the analysis

```{r description, include=FALSE}

summary(ca_county_demographic)

summary(ca_county_mortality)

summary(hcai_healthcare_construction)

```
