---
title: "Group S: Milestone #2"
author: "PhuongTseng & Carri Beshears"
date: '2022-10-02'
output: pdf_document
---

This is a team assignment; each team should complete and turn in a PDF created 
from an Rmd via Github. Please include code and output for the following components:

## Description of dataset
What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
How does the data set relate to the group problem statement and question?

ANSWER: This project will use 3 data sets to identify 5 California counties for additional funding for healthcare facility improvement. First, information from the 2012 California county census will be reviewed to to isolate counties in rural areas, with residents of a high median age, and with a high proportion of renters over homeowners. Next, mortaility surveillance data will be used to determine mortality from chronic conditions will be aggregated by county for the years 2014-2020. Finally, HCAI data will be used to account for current healthcare funding by county from 2014-2020. All of these data sets are avaliable from the CA.gov open data portal. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(data.table)
library(dplyr)
library(readr)

path <- getwd()
setwd(path)

#import demographic data
ca_county_demographic <- fread("./ca_county_demographic.csv")
#import mortality data
ca_county_mortality <- fread("./ca_county_mortality.csv")
#import healthcare construction data with fread
hcai_healthcare_construction <- fread("./hcai_healthcare_construction.csv")
```

## Import statement
NOTE: Please use datasets available in the PHW251 Project Data github repoLinks to an external site. 
(This is important to make sure everyone is using the same datasets)

Use appropriate import function and package based on the type of file
Utilize function arguments to control relevant components 
(i.e. change column types, column names, missing values, etc.)

#### Document the import process
```{r import, include=FALSE}
#use dplyr::select specific columns
demog <- ca_county_demographic %>% dplyr::select(name, pop2012, pop12_sqmi, med_age, owner_occ, renter_occ)
#check first 2 records
head(demog, 2)

#use dplyr::rename_with to make column names lower case
mortality <- ca_county_mortality %>% dplyr::rename_with(~ tolower(gsub(" ", "_", .x, fixed = TRUE)))

head(mortality, 2)

#same as above and substitute a dash "_" for space
construction <- hcai_healthcare_construction %>% dplyr::rename_with(~ tolower(gsub(" ", "_", .x, fixed = TRUE)))
#see the last 2 records
tail(construction, 2)
```


\newpage

#### Identify data types for 5+ data elements/columns/variables
- Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
- Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
- Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type?

ANSWER: The 5 elements that will be used for this scenario start with 2 existing variables showing population per square mile (pop12_sqmi) and median age for the 2012 California county census (med_age). 

```{r data elements, include=FALSE}
typeof(ca_county_demographic$pop12_sqmi)
typeof(ca_county_demographic$med_age)
typeof(ca_county_demographic$owner_occ)
typeof(ca_county_demographic$renter_occ)
typeof(ca_county_mortality$Count)
typeof(hcai_healthcare_construction$`Total Costs of OSHPD Projects`)



```

An additional variable will be created from 2 existing variables (owner_occ and renter_occ) to show the proportion of renters to homeowners. These are both currently numeric type variables and will not require a conversion. 2 additional variables will be added showing the county level mortality counts and project funding from raw data. These are both current stored at character type variables and will need to be converted to numeric type to aggregate and allow for analysis. 

\newpage

#### Provide a basic description of the 5+ data elements
- Numeric: mean, median, range
- Character:  unique values/categories
- Or any other descriptive that will be useful to the analysis

```{r description, include=FALSE}

summary(ca_county_demographic$pop12_sqmi)
summary(ca_county_demographic$med_age)
summary(ca_county_demographic$owner_occ)
summary(ca_county_demographic$renter_occ)

```

Other required elements that will be used for analysis includes the aggregated data that will be created to determine county level information on mortality counts and project funding. These items are currently stored as character variables and will need to be converted prior to working with and moving them onto the main data set. 
