---
title: "GroupS_Milestone3"
author: "PhuongTseng"
date: '2022-11-07'
output: pdf_document
---


## Clean dataset with descriptive statistics for relevant data elements

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

ca_county_demographic <- fread("./ca_county_demographic.csv")


```

#### Subset rows or columns, as needed
- Create new variables needed for analysis (minimum 2)
  - New variables should be created based on existing columns; for example
    - Calculating a rate
    - Combining character strings
  - If no new values are needed for final tables/graphs, please create 2 new variables anyway
  
Using our 3 vectors:  
1) ca_county_demographic  
2) ca_county_mortality  
3) hcai_healthcare_construction  
  
```{r subset, eval=FALSE, include=FALSE}

ca_county_demographic

ca_county_mortality

hcai_healthcare_construction



```

#### Clean variables needed for analysis (minimum 2)
  - Examples
    - Recode invalid values
    - Handle missing fields
    - Recode categories
  - If not needed for final analysis, please create at least 2 new variables anyway
  
```{r}


```

#### Data dictionary based on clean dataset (minimum 4 data elements), including:
  - Variable name
  - Data type
  - Description
  
```{r}


```

#### One or more tables with descriptive statistics for 4 data element
```{r}


```

#### PDF that is professionally prepared for presentation
  - Each part of the milestone is clearly on one page (use \newpage to push to a new page)
  - Only the necessary information is outputted (you should suppress, for example, entire data frame outputs)
  - Use of headers and sub headers to create an organized document

```{r}


```

